Manuscript 1 code

### 2024-09-13 manuscript_1_code.Rmd
Wrote models for quad and site evenness, made figure (proposed 3b) with quad-level evenness versus cover :cow2:

```{r COME BACK HERE set up evenness}
site_evenness_df <- alpha_diversity_site_macro %>% 
  drop_na(evenness) %>%
  group_by(year, habitat) %>% 
  summarise(eveness_mean = mean(evenness),
            eveness_se = se(evenness))

@@

## set up data for model

```{r site data setup}
alpha_diversity_site_macro_with_cover <- alpha_diversity_site_macro %>% 
  mutate(percent_cover = rowSums(site_macro[,-which(names(site_macro)  %in% colnames(site_meta_df))])) %>% 
  mutate(prop_cover = percent_cover / 100) %>% 
  select(-percent_cover) %>% 
  mutate(cover_trans = sv_trans(prop_cover))
```

@@

### quad

```{r figure 3a model}
alpha_diversity_quad_macro_with_cover$cover_trans <- sv_trans(alpha_diversity_quad_macro_with_cover$macroalgal_prop_cover)

@@

### site
```{r site mod}
site_cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover)
summary(site_cover_mod)
Anova(site_cover_mod)
#                    Chisq Df Pr(>Chisq)    
# richness         58.9195  1  1.643e-14 ***
# habitat           6.6247  3    0.08487 .  
# richness:habitat  3.2967  3    0.34811    
hist(residuals(site_cover_mod)) # some positive skew, not the worst thing ever
plot(residuals(site_cover_mod) ~ fitted(site_cover_mod)) # a little heteroscedastic but can't be less than 0
r.squaredGLMM(site_cover_mod) # 0.2212534, 0.5990825
emtrends(site_cover_mod, pairwise ~ habitat, var = "richness") # no difference between the habitats
```

## figure 3a plot setup

```{r figure 3a setup}
diversity_ranges_quad <- alpha_diversity_quad_macro_with_cover %>%
  group_by(habitat) %>%
  summarise(min_richness = min(richness),
            max_richness = max(richness),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

@@

## figure 3a plot

@@

## Figure 3d

>>>>

```{r set up evenness}
site_evenness_df <- alpha_diversity_site_macro %>% 
  drop_na(evenness) %>%
  group_by(year, habitat) %>% 
  summarise(eveness_mean = mean(evenness),
            eveness_se = se(evenness))

@@

## set up site-level data for model
```{r site cover by diversity data setup}
# alpha_diversity_site_macro_with_cover <- alpha_diversity_site_macro %>% 
#   dplyr::mutate(percent_cover = rowSums(site_macro[,-which(names(site_macro)  %in% colnames(site_meta_df))])) %>% 
#   mutate(prop_cover = percent_cover / 100) %>% 
#   select(-percent_cover) %>% 
#   mutate(cover_trans = sv_trans(prop_cover))

#there is no reasonable reason why the above chunk doesn't work, but it doesn't, so we did it this way
alpha_diversity_site_macro$percent_cover <- rowSums(site_macro[,-which(names(site_macro)  %in% colnames(site_meta_df))])

alpha_diversity_site_macro_with_cover <- alpha_diversity_site_macro %>% 
  dplyr::mutate(prop_cover = percent_cover / 100) %>%
  select(-percent_cover)

alpha_diversity_site_macro_with_cover$cover_trans <- sv_trans(alpha_diversity_site_macro_with_cover$prop_cover)
```

@@

### quad
```{r figure 3a richness quad model}
alpha_diversity_quad_macro_with_cover$cover_trans <- sv_trans(alpha_diversity_quad_macro_with_cover$macroalgal_prop_cover)

@@

```{r 3b evenness quad level model}
alpha_diversity_quad_macro_with_cover_evenness_mod <- alpha_diversity_quad_macro_with_cover %>% 
  drop_na(evenness)

evenness_cover_mod_quad <- glmmTMB(cover_trans ~ evenness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro_with_cover_evenness_mod)
summary(evenness_cover_mod_quad)
Anova(evenness_cover_mod_quad)
#                     Chisq Df Pr(>Chisq)    
# richness         30916.73  1  < 2.2e-16 ***
# habitat            141.78  3  < 2.2e-16 ***
# richness:habitat   294.54  3  < 2.2e-16 ***
hist(residuals(evenness_cover_mod_quad)) # slight pos skeq
plot(residuals(evenness_cover_mod_quad) ~ fitted(evenness_cover_mod_quad)) # negative trend but it can't be less than 0 so not too concerned. some wave pattern
r.squaredGLMM(evenness_cover_mod_quad) # 0.1749907, 0.4042826
emtrends(evenness_cover_mod_quad, pairwise ~ habitat, var = "evenness") #backreef is different from everything except outer 17, but just barely. everyone else is same same xoxo

### site richness
```{r site richness mod}
site_cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover)
summary(site_cover_mod)
Anova(site_cover_mod)
#                    Chisq Df Pr(>Chisq)    
# richness         58.9195  1  1.643e-14 ***
# habitat           6.6247  3    0.08487 .  
# richness:habitat  3.2967  3    0.34811    
hist(residuals(site_cover_mod)) # some positive skew, not the worst thing ever
plot(residuals(site_cover_mod) ~ fitted(site_cover_mod)) # a little heteroscedastic but can't be less than 0
r.squaredGLMM(site_cover_mod) # 0.2212534, 0.5990825
emtrends(site_cover_mod, pairwise ~ habitat, var = "richness") # no difference between the habitats
```

### site evenness 
```{r}
alpha_diversity_site_macro_with_cover_evenness_mod <- alpha_diversity_site_macro_with_cover %>% 
  drop_na(evenness)

evenness_site_cover_mod <- glmmTMB(cover_trans ~ evenness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover_evenness_mod)
summary(evenness_site_cover_mod)
Anova(evenness_site_cover_mod)
#                    Chisq Df Pr(>Chisq)    
# richness         58.9195  1  1.643e-14 ***
# habitat           6.6247  3    0.08487 .  
# richness:habitat  3.2967  3    0.34811    
hist(residuals(evenness_site_cover_mod)) # some positive skew, not the worst thing ever
plot(residuals(evenness_site_cover_mod) ~ fitted(evenness_site_cover_mod)) # a little heteroscedastic but can't be less than 0
r.squaredGLMM(evenness_site_cover_mod) #0.1206569, 0.5852773
emtrends(evenness_site_cover_mod, pairwise ~ habitat, var = "evenness") # only the outers are different from each other
```

## figure 3a plot setup
```{r figure 3a setup}
diversity_ranges_quad <- alpha_diversity_quad_macro_with_cover %>%
  group_by(habitat) %>%
  summarise(min_richness = min(richness),
            max_richness = max(richness),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

@@

## figure 3a quad level richness versus cover

@@

## Figure 3B quad level evenness versus cover
```{r COME BACK HERE figure 3b setup}
diversity_ranges_quad_evenness <- alpha_diversity_quad_macro_with_cover_evenness_mod %>%
  group_by(habitat) %>%
  summarise(min_evenness = min(evenness),
            max_evenness = max(evenness),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

cover_emm_evenness <- emmip(evenness_cover_mod_quad,
                        habitat ~ evenness,
                        at = list(evenness = seq(0,10,1)), plotit = F, CIs = T)

filtered_cover_effects_evenness <- cover_emm %>%
  left_join(diversity_ranges_quad_evenness, by = c("habitat")) %>%
  filter(evenness >= min_evenness & evenness <= max_evenness)

alpha_diversity_quad_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_quad_macro_with_cover_evenness_mod$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

filtered_cover_effects_evenness$habitat <- factor(filtered_cover_effects_evenness$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

```{r}
(figure_3b <- 
  ggplot() +
  geom_point(data = alpha_diversity_quad_macro_with_cover_evenness_mod,
             aes(x = evenness, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
  geom_line(data = filtered_cover_effects_evenness, aes(x = evenness, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects_evenness, aes(x = evenness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Cover\n", x = "Macroalgal evenness", title = "3a.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

#ggsave(filename = "output/figure_3b.png", figure_3b, height = 10, width = 14)
```

## Figure 3E site level richness versus cover
